Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.tex
Commit 2024-06-09 22:17

According to this study, in the period 1997 -- 2005, the mean AOD at
\(340\,\text{nm}\) \(0.403\) with a trend of \(-3.8±0.93\,\%\) per year,
corresponding to a change of \(0.0153\) per year. Using an Ångström
coefficient \(\alpha = 1.6\), this translates to a change in the
Ångström coefficient \(\beta=0.00272\) per year (or \(\beta=0.084\) in
1997 and \(\beta=0.059\) in 2005).

Simulations with uvspec for the above Ångström coefficients, with WC of
\(`r round(BR[year==1997, pw_avg_mm],1)`\,\text{mm}\) and
\(15.4\,\text{mm}\) for 1997 and 2005 respectively and for a SZA of
\(55^\circ\)

By interpolating the resulting clear sky irradiance, we have a trend of
\(+0.21\,\%\) per year for the clear sky irradiance. For the period 2005
-- 2005 we used the mean monthly values from AERONET, with a similar
simulation scheme. We computed the trend of clear sky radiation for each
month of the year, which gave us a mean clear sky radiation trend of
\(+0.14\,\%\) per year. We applied this long term change to the clear
sky irradiance, in order to create a more realistic representation of
the irradiance for the whole study period
(Figure\nobreakspace{}\ref{fig:CS-change}).

@@

\caption{Simulated long term change of the clear sky irradiance relevant to its climatological values, due to changed of AOD.}\label{fig:CS-change}
\end{figure}

\hypertarget{ce-criteria-investigation}{%
\subsection{CE Criteria investigation}\label{ce-criteria-investigation}}

Our main focus was to positively identify over irradiance events related
to CEs. For this reason we had to select an appropriate set of
atmospheric parameterization, in Libradtran, and define an relevant
threshold for the CE identification. The selection of the atmospheric
parameterization will provide us with a reference, that has a good
agreement with the clear sky conditions. Additional, with the
implementation of the long term change of AOD, as described above, it
will be able to align well with most of the natural variation of the
physical quantity. The selection of adequate threshold, will allow us to
have a robust method to distinguish the CE cases, which will compensate
for the limited accuracy of the input data, and the natural variability
of the atmosphere.

First, we evaluated the performance of the modelled clear radiation, for
each of the atmospheric level of clearness we have created as reference,
in relation to the measured GHI data, and choose as a representative of
the clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki, with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the upper limit of CE identification, we
used a value relative to clear sky irradiance (\(4\,\%\)) with an
additional constant offset of \(15\,\text{W}/\text{m}^2\), as described
in Equation\nobreakspace\ref{eq:CE4}). \begin{equation}
\text{CE} : \text{E}_\text{i} > 15 + 1.04 \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation} where: \(\text{E}\) the measured irradiance,
\(\text{E}_\text{CSm}\) the selected modelled clear sky irradiance, and
\(i\) each of the one-minute observation. This is the criterion of our
CE identification.

>>>>

According to this study, in the period 1997 -- 2005, the mean AOD at
\(340\,\text{nm}\) \(0.403\) with a trend of \(-3.8\pm0.93\,\%\) per
year, corresponding to a change of \(0.0153\) per year. Using an
Ångström coefficient \(\alpha = 1.6\), this translates to a change in
the Ångström coefficient \(\beta=0.00272\) per year (or \(\beta=0.084\)
in 1997 and \(\beta=0.059\) in 2005). Simulations with uvspec for the
above Ångström coefficients, with WC of \(15.7\,\text{mm}\) and
\(15.4\,\text{mm}\) for 1997 and 2005 respectively and for a SZA of
\(55^\circ\) reveal trend of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean of clear
sky irradiance, and finally the trend of \(+0.14\,\%\) per year. We
applied these two long term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

\caption{Simulated long term change in clear sky irradiance relative to the climatological values due to changes in AOD.}\label{fig:CS-change}
\end{figure}

\hypertarget{criteria-for-the-identification-of-ce-events}{%
\subsection{Criteria for the identification of CE
events}\label{criteria-for-the-identification-of-ce-events}}

In this study our main focus was to quantify over irradiance events
related to CEs. A key issue for achieving this goal is to define a
threshold for the CE identification, representative of the clear-sky
irradiance at the time of each GHI measurement. This depends on the
selection of the appropriate atmospheric parameterization for the RTM
simulations. The implementation of the long term change of AOD,
discussed in section\nobreakspace{}\ref{rtmcs}, allows capturing a large
part of the natural variability of clear-sky GHI. However, the
short-term variability of AOD cannot be taken adequately into account
when using monthly values for the model simulations. We tried different
approaches to increase the robustness of the methodology and compensate
for the limited accuracy of the RTM input data and the unpredictable
natural variability of the atmosphere.

First, we evaluated the performance of the modelled
\(\text{GHI}_\text{ref}\) in relation to the measured GHI for different
levels of atmospheric clearness, by using in the RTM the monthly
climatological AOD and WC, less their respective standard deviations.
the atmospheric level of clearness we have created as reference, in
relation to the measured GHI data, and choose as a representative of the
clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki, with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the upper limit of CE identification, we
used a value relative to clear sky irradiance (\(4\,\%\)) with an
additional constant offset of \(15\,\text{W}/\text{m}^2\), as described
in Equation\nobreakspace\ref{eq:CE4}). \begin{equation}
\text{CE} : \text{E}_\text{i} > 15 + 1.04 \cdot \text{E}_\text{CSm,i} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation} where: \(\text{E}\) the measured irradiance,
\(\text{E}_\text{CSm}\) the selected modelled clear sky irradiance, and
\(i\) each of the one-minute observation. This is the criterion of our
CE identification.
